---
title: "NHTS Descriptive Statistics_Person_CGRiffeth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Loading my libraries.

```{r libraries}
library(tidyverse)
library(survey)
library(srvyr)
library(treemapify)
library(knitr)
```

## Loading NHTS Data - Person

Loading variables from the NHTS household person characteristics file.
One record per household person.

The following are continuous rather than categorical variables: PTUSED, R_AGE.
A mutation has not been written for HHSTATE.

* BORNINUS = Born in United States
* CENSUS_D = 2010 Census division classification for the respondent's home address
* HBHTNRNT = Category of the percent of renter-occupied housing in the census block group of the household's home location
* HHSIZE = Count of household members
* HHSTATE = Household state
* HOMEOWN = Home Ownership
* PTUSED = Count of Public Transit Usage
* R_AGE = Age
* R_SEX = Gender
* WRKTRANS = Mode to Work
* WRK_HOME = Work from Home

```{r NHTS Data}
hh <- read_csv("NHTS_data/perpub.csv") %>%
  select(HOUSEID, PERSONID, BORNINUS, CENSUS_D, 
         HBHTNRNT, HHSIZE, HHSTATE, HOMEOWN, 
         PTUSED, R_AGE, R_SEX, WRKTRANS, WRK_HOME) %>%
  mutate(region = case_when(CENSUS_D == "01" ~ "New England",
                            CENSUS_D == "02" ~ "Middle Atlantic",
                            CENSUS_D == "03" ~ "East North Central",
                            CENSUS_D == "04" ~ "West North Central",
                            CENSUS_D == "05" ~ "South Atlantic",
                            CENSUS_D == "06" ~ "East South Central",
                            CENSUS_D == "07" ~ "West South Central",
                            CENSUS_D == "08" ~ "Mountain",
                            CENSUS_D == "09" ~ "Pacific",
                            TRUE ~ "Other")) %>%
  mutate(rent_hh = case_when(HBHTNRNT == "0" ~ 2,
                             HBHTNRNT == "05" ~ 10,
                             HBHTNRNT == "20" ~ 20,
                             HBHTNRNT == "30" ~ 30,
                             HBHTNRNT == "40" ~ 40,
                             HBHTNRNT == "50" ~ 50,
                             HBHTNRNT == "60" ~ 60,
                             HBHTNRNT == "70" ~ 70,
                             HBHTNRNT == "80" ~ 80,
                             HBHTNRNT == "90" ~ 90,
                             HBHTNRNT == "95" ~ 97,
                             TRUE ~ -1)) %>%
  mutate(hh_size = case_when(HHSIZE == "1" ~ 1,
                             HHSIZE == "2" ~ 2,
                             HHSIZE == "3" ~ 3,
                             HHSIZE == "4" ~ 4,
                             HHSIZE == "5" ~ 5,
                             HHSIZE == "6" ~ 6,
                             HHSIZE == "7" ~ 7,
                             HHSIZE == "8" ~ 8,
                             HHSIZE == "9" ~ 9,
                             HHSIZE == "10" ~ 10,
                             HHSIZE == "11" ~ 11,
                             HHSIZE == "12" ~ 12,
                             HHSIZE == "13" ~ 13,
                             TRUE ~ -1)) %>%
  mutate(own_hh = case_when(HOMEOWN == "01" ~ "own",
                            HOMEOWN == "02" ~ "rent",
                            TRUE ~ "Other")) %>%
  mutate(gender = case_when(R_SEX == "01" ~ "Male",
                            R_SEX == "02" ~ "Female",
                            TRUE ~ "Other")) %>%
  mutate(modeshare = case_when(WRKTRANS == "01" ~ "Walk",
                               WRKTRANS == "02" ~ "Bicyle",
                               WRKTRANS == "03" ~ "Car",
                               WRKTRANS == "04" ~ "SUV",
                               WRKTRANS == "05" ~ "Van",
                               WRKTRANS == "06" ~ "Pickup Truck",
                               WRKTRANS == "07" ~ "Golf car / Segway",
                               WRKTRANS == "08" ~ "Motorcycle / Moped",
                               WRKTRANS == "09" ~ "RV",
                               WRKTRANS == "10" ~ "School bus",
                               WRKTRANS == "11" ~ "Public or commuter bus",
                               WRKTRANS == "12" ~ "Paratransit",
                               WRKTRANS == "13" ~ "Private / Charter / Tour / Shuttle Bus",
                               WRKTRANS == "14" ~ "City-to-city bus",
                               WRKTRANS == "15" ~ "Amtrak / Commuter Rail",
                               WRKTRANS == "16" ~ "Subway / Elevated / Ligtht Rail / Street Car",
                               WRKTRANS == "17" ~ "Taxi / Limo / Rideshare",
                               WRKTRANS == "18" ~ "Rental car",
                               WRKTRANS == "19" ~ "Airplane",
                               WRKTRANS == "20" ~ "Boat / Ferry / Water taxi",
                               TRUE ~ "Other")) %>%
  mutate(wfh = case_when(WRK_HOME == "01" ~ "Yes",
                         WRK_HOME == "02" ~ "No",
                         TRUE ~ "Other")) 
```
##Survey Weights

```{r Create survey}
hh_svy <- hh %>%
  as_survey(HOUSEID, weight = WTHHFIN)
  as_survey(HHRESP, weight = WTPERFIN)
  as_survey(TDCASEID, weight = WTTRDFIN)
```