---
title: "NHTS Descriptive Statistics_Trip_CGRiffeth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Loading my libraries.

```{r libraries}
library(tidyverse)
library(survey)
library(srvyr)
library(treemapify)
library(knitr)
```

## Loading NHTS Data - Trip

Loading variables from the NHTS household person trip characteristics file.
One record per household person’s travel day trip, if at least one trip made.

The following are continuous rather than categorical variables: DWELTIME, R_AGE, TRVLCMIN, TRWAITTM.

* CENSUS_D = 2010 Census division classification for the respondent's home address
* DBHTNRNT = Category of the percent of renter-occupied housing in the census block group of the trip destination's location
* DBHUR = Urban / Rural indicator - Trip Destination Block group
* DWELTIME = Time at destination
* HHSIZE = Count of household members
* HHSTATE = Household state
* HOMEOWN = Home Ownership
* OBHTNRNT = Category of the percent of renter-occupied housing in the census block group of the trip origin's location
* OBRESDN = Category of housing units per square mile in the census block group of the trip origin's location
* PUBTRANS = Public transportation used on trip
* R_AGE = Age
* R_SEX = Gender
* TDWKND = Weekend trip
* TRVLCMIN = Trip Duration in Minutes
* TRWAITTM = Transit wait time in minutes
* VEHTYPE = Vehicle Type

```{r NHTS Data}
trip <- read_csv("NHTS_data/trippub.csv") %>%
  select(TDTRPNUM, CENSUS_D, DBHTNRNT, DBHUR, 
         DWELTIME, HHSIZE, HHSTATE, HOMEOWN, 
         OBHTNRNT, OBRESDN, PUBTRANS, R_AGE, 
         R_SEX, TDWKND, TRVLCMIN, TRWAITTM, VEHTYPE) %>%
  mutate(region = case_when(CENSUS_D == "01" ~ "New England",
                            CENSUS_D == "02" ~ "Middle Atlantic",
                            CENSUS_D == "03" ~ "East North Central",
                            CENSUS_D == "04" ~ "West North Central",
                            CENSUS_D == "05" ~ "South Atlantic",
                            CENSUS_D == "06" ~ "East South Central",
                            CENSUS_D == "07" ~ "West South Central",
                            CENSUS_D == "08" ~ "Mountain",
                            CENSUS_D == "09" ~ "Pacific",
                            TRUE ~ "Other")) %>%
  mutate(rent_destination = case_when(DBHTNRNT == "0" ~ 2,
                                      DBHTNRNT == "05" ~ 10,
                                      DBHTNRNT == "20" ~ 20,
                                      DBHTNRNT == "30" ~ 30,
                                      DBHTNRNT == "40" ~ 40,
                                      DBHTNRNT == "50" ~ 50,
                                      DBHTNRNT == "60" ~ 60,
                                      DBHTNRNT == "70" ~ 70,
                                      DBHTNRNT == "80" ~ 80,
                                      DBHTNRNT == "90" ~ 90,
                                      DBHTNRNT == "95" ~ 97,
                                      TRUE ~ -1)) %>%
  mutate(dest_den = case_when(DBHUR == "C" ~ "Second City",
                              DBHUR == "R" ~ "Rural",
                              DBHUR == "S" ~ "Suburban",
                              DBHUR == "T" ~ "Small Town",
                              DBHUR == "U" ~ "Urban",
                              TRUE ~ "Other")) %>%
  mutate(hh_size = case_when(HHSIZE == "1" ~ 1,
                             HHSIZE == "2" ~ 2,
                             HHSIZE == "3" ~ 3,
                             HHSIZE == "4" ~ 4,
                             HHSIZE == "5" ~ 5,
                             HHSIZE == "6" ~ 6,
                             HHSIZE == "7" ~ 7,
                             HHSIZE == "8" ~ 8,
                             HHSIZE == "9" ~ 9,
                             HHSIZE == "10" ~ 10,
                             HHSIZE == "11" ~ 11,
                             HHSIZE == "12" ~ 12,
                             HHSIZE == "13" ~ 13,
                             TRUE ~ -1)) %>%
  mutate(own_hh = case_when(HOMEOWN == "01" ~ "own",
                            HOMEOWN == "02" ~ "rent",
                            TRUE ~ "Other")) %>%
  mutate(rent_trip_origin = case_when(OBHTNRNT == "0" ~ 2,
                                      OBHTNRNT == "05" ~ 10,
                                      OBHTNRNT == "20" ~ 20,
                                      OBHTNRNT == "30" ~ 30,
                                      OBHTNRNT == "40" ~ 40,
                                      OBHTNRNT == "50" ~ 50,
                                      OBHTNRNT == "60" ~ 60,
                                      OBHTNRNT == "70" ~ 70,
                                      OBHTNRNT == "80" ~ 80,
                                      OBHTNRNT == "90" ~ 90,
                                      OBHTNRNT == "95" ~ 97,
                                      TRUE ~ -1)) %>%
  mutate(house_density = case_when(OBRESDN == "50" ~ 2,
                                   OBRESDN == "300" ~ 10,
                                   OBRESDN == "750" ~ 20,
                                   OBRESDN == "1500" ~ 30,
                                   OBRESDN == "3000" ~ 40,
                                   OBRESDN == "7000" ~ 50,
                                   OBRESDN == "17000" ~ 60,
                                   OBRESDN == "30000" ~ 70,
                                   OBRESDN == "80" ~ 80,
                                   OBRESDN == "90" ~ 90,
                                   OBRESDN == "95" ~ 97,
                                   TRUE ~ -1)) %>%
  mutate(pub_transit = case_when(PUBTRANS == "01" ~ "Yes",
                                 PUBTRANS == "02" ~ "No",
                                 TRUE ~ "Other")) %>%
  mutate(gender = case_when(R_SEX == "01" ~ "Male",
                            R_SEX == "02" ~ "Female",
                            TRUE ~ "Other")) %>%
  mutate(weekend_trips = case_when(TDWKND == "01" ~ "own",
                                    TDWKND == "02" ~ "rent",
                                    TRUE ~ "Other")) %>%
  mutate(veh_type = case_when(VEHTYPE == "01" ~ "Automobile/Car/Station Wagon",
                              VEHTYPE == "02" ~ "Van (Mini/Cargo/Passenger)",
                              VEHTYPE == "03" ~ "SUV",
                              VEHTYPE == "04" ~ "Pickup Truck",
                              VEHTYPE == "05" ~ "Other Truck",
                              VEHTYPE == "06" ~ "RV",
                              VEHTYPE == "07" ~ "Motorcycle",
                              TRUE ~ "Other"))
```
##Survey Weights

```{r Create survey}
trip_svy <- trip %>%
  as_survey(HOUSEID, weight = WTHHFIN)
  as_survey(HHRESP, weight = WTPERFIN)
  as_survey(TDCASEID, weight = WTTRDFIN)
```
