---
title: "NHTS Descriptive Statistics_HH_CGRiffeth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Loading my libraries.

```{r libraries}
library(tidyverse)
library(survey)
library(srvyr)
library(treemapify)
library(knitr)
```

## Loading NHTS Data - Household

Loading variables from the NHTS household characteristics file.
One record per household unit.

* CAR = Frequency of Personal Vehicle Use for Travel
* CENSUS_D = 2010 Census division classification for the respondent's home address
* HBHTNRNT = Category of the percent of renter-occupied housing in the census block group of the household's home location
* HHSIZE = Count of household members
* HHSTATE = Household state
* HOMEOWN = Home Ownership
* PARA = Frequency of Paratransit Use for Travel
* PLACE = Travel is a Financial Burden
* PRICE = Price of Gasoline Affects Travel.

```{r NHTS Data}
pers <- read_csv("NHTS_data/hhpub.csv") %>%
  select(HOUSEID, CAR, CENSUS_D, HBHTNRNT, HOMEOWN, PARA, PLACE, PRICE) %>%
  mutate(car_use = case_when(CAR == "01" ~ "Daily",
                             CAR == "02" ~ "A Few Times a Week",
                             CAR == "03" ~ "A Few Times a Month",
                             CAR == "04" ~ "A Few Times a Year",
                             CAR == "05" ~ "Never",
                             TRUE ~ "Other")) %>%
  mutate(region = case_when(CENSUS_D == "01" ~ "New England",
                            CENSUS_D == "02" ~ "Middle Atlantic",
                            CENSUS_D == "03" ~ "East North Central",
                            CENSUS_D == "04" ~ "West North Central",
                            CENSUS_D == "05" ~ "South Atlantic",
                            CENSUS_D == "06" ~ "East South Central",
                            CENSUS_D == "07" ~ "West South Central",
                            CENSUS_D == "08" ~ "Mountain",
                            CENSUS_D == "09" ~ "Pacific",
                            TRUE ~ "Other")) %>%
  mutate(rent_hh = case_when(HBHTNRNT == "0" ~ 0,
                             HBHTNRNT == "05" ~ 10,
                             HBHTNRNT == "20" ~ 20,
                             HBHTNRNT == "30" ~ 30,
                             HBHTNRNT == "40" ~ 40,
                             HBHTNRNT == "50" ~ 50,
                             HBHTNRNT == "60" ~ 60,
                             HBHTNRNT == "70" ~ 70,
                             HBHTNRNT == "80" ~ 80,
                             HBHTNRNT == "90" ~ 90,
                             HBHTNRNT == "95" ~ 97,
                             TRUE ~ -1)) %>%
  mutate(own_hh = case_when(HOMEOWN == "01" ~ "Own",
                            HOMEOWN == "02" ~ "Rent",
                            TRUE ~ -"Other")) %>%
  mutate(paratransit_use = case_when(PARA == "01" ~ "Daily",
                                     PARA == "02" ~ "A Few Times a Week",
                                     PARA == "03" ~ "A Few Times a Month",
                                     PARA == "04" ~ "A Few Times a Year",
                                     PARA == "05" ~ "Never",
                                     TRUE ~ "Other")) %>%
  mutate(finan_burden = case_when(PLACE == "01" ~ "Strongly Agree",
                                  PLACE == "02" ~ "Agree",
                                  PLACE == "03" ~ "Neither Agree nor Disagree",
                                  PLACE == "04" ~ "Disagree",
                                  PLACE == "05" ~ "Strongly Disagree",
                                  TRUE ~ "Other")) %>%
  mutate(fuel_burden = case_when(PRICE == "01" ~ "Strongly Agree",
                                 PRICE == "02" ~ "Agree",
                                 PRICE == "03" ~ "Neither Agree nor Disagree",
                                 PRICE == "04" ~ "Disagree",
                                 PRICE == "05" ~ "Strongly Disagree",
                                 TRUE ~ "Other"))
rlang::last_error()
```

## Survey Weights

``` {r survey weights}
pers_svy <- pers %>%
  as_survey(HOUSEID, weight = WTHHFIN)
  as_survey(HHRESP, weight = WTPERFIN)
  as_survey(TDCASEID, weight = WTTRDFIN)
```

## Descriptive Statistics

``` {r descriptstats}
ev_shares_list <- svytable(~type, vehs_svy)

ev_shares_tibble <- tibble(`Type of vehicle` = 
                             factor(names(ev_shares_list),
                                    levels = c("Gas",
                                               "Diesel",
                                               "Biodiesel",
                                               "Hybrid",
                                               "Plug-in hybrid",
                                               "Fully electric",
                                               "Other")), 
                           `Number of vehicles` =
                                as.numeric(round(ev_shares_list,0)),
                           `Percent of total` = paste(
                             round((ev_shares_list / 
                                      sum(ev_shares_list)) * 100, 2),
                             "%", sep = "")) %>%
  arrange(`Type of vehicle`)

kable(ev_shares_tibble, 
      format.args = list(big.mark = ","),
      align = "lrr")
```

